---
title: "CUNY SPS - Master of Science in Data Science - DATA607 "
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assignment 2: Database Connection With R
# Mario Pena
# 9/8/2019

In this assignment I have chosen six of the most popular Hollywood movies, and I have asked 6 people to rate each of the movies they've seen on a scale from 1 (being terrible) to 5 (being great).

I have stored the observations in a SQL database, so I will first need to load this data into an R dataframe.

We'll need to load the following libraries:
```{r}
library(RMySQL)
library(dbConnect)
```
```{r, echo=FALSE}
options(database_password = 'test123*')
```
And connect to MySQL server:
```{r}
con = dbConnect (MySQL(), user = 'root', password = getOption('database_password'), dbname = 'movie_ratings', Host = 'localhost')
```
Now I can see a list of the tables in my "movie_ratings" database:
```{r}
dbListTables(con)
```
As you can see, I have created 3 different tables, each stores information regarding audience, movies and ratings.

I will create 3 queries within R using these tables:
```{r}
myquery1 <- "select * from audience;"
myquery2 <- "select * from movies;"
myquery3 <- "select * from ratings;"
```
I will save them as dataframes in this next step and display their content:
```{r}
audience <- dbGetQuery(con, myquery1)
audience
movies <- dbGetQuery(con, myquery2)
movies
ratings <- dbGetQuery(con, myquery3)
ratings
```
I will merge 'movies' and 'ratings' so we can have a clearer picture of how each movie was rated:
```{r}
movie_ratings <- merge(movies, ratings, by.x="movieid")
movie_ratings
```
Let's take a look at what the summmary of this new table looks like:
```{r}
summary(movie_ratings)
```
We can see that the average rating for all the movies in this database is 4.212. This tells us these movies were very well received by the audience who watched them.

Now let's see what is the average movie rating per movie:
```{r}
avg_movie_rating <- aggregate(x=movie_ratings["rating"], by = list(movie=movie_ratings$title), FUN = mean, na.rm=TRUE)
avg_movie_rating
```
We can see that "The Lion King (Original)" has the highest average rating in this list, while "Jaws" has the lowest average rating as also shown in the barplot below:
```{r}
library(ggplot2)
ggplot(aes(x = movie, y = rating), data = avg_movie_rating) + geom_bar(stat = "identity")
```
